### 2015-12-09 abstract.tex
typos and small changes

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to norms of loss vectors by scale
invariance, i.e., our algorithms make exactly the same decisions if the
sequence of loss vectors is multiplied by any positive constant.  Our
algorithms work for any decision set, bounded or unbounded.  For unbounded
decisions sets, these are the first adaptive algorithms for online linear
optimization.

>>>>

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to the norms of the loss vectors by scale
invariance, i.e., our algorithms make exactly the same decisions if the
sequence of loss vectors is multiplied by any positive constant.  Our
algorithms work for any decision set, bounded or unbounded.  For unbounded
decisions sets, these are the first adaptive algorithms for online linear
optimization with a non-vacuous bound.

### 2015-12-09 conclusions.tex
typos and small changes

Similar, but weaker result holds for a scale-free algorithm based on
\textsc{Mirror Descent}. However, we have also shown this algorithm is strictly
weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
we gave examples of regularizers for which the associated Bregman divergence is
unounded and the scale-fre version of \textsc{Mirror Descent} has $\Omega(T)$
regret or worse.

>>>>

Similar, but weaker result holds for a scale-free algorithm based on
\textsc{Mirror Descent}. However, we have also shown this algorithm is strictly
weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
we gave examples of regularizers for which the associated Bregman divergence is
unounded and the scale-free version of \textsc{Mirror Descent} has $\Omega(T)$
regret or worse.

### 2015-12-09 introduction.tex
typos and small changes

Online Linear Optimization (OLO) is a problem where an algorithm repeatedly
chooses a point $w_t$ from a convex decision set $K$, observes an arbitrary, or
even adversarially chosen, loss vector $\ell_t$ and suffers loss $\langle
\ell_t, w_t \rangle$.  The goal of the algorithm is to have a small cumulative
loss. Performance of an algorithm is evaluated by the so-called regret, which
is the difference of cumulative losses of the algorithm and of the
(hypothetical) strategy that would choose in every round the same best point in
hindsight.

@@

Rather than trying to design strategies for a particular form of loss vectors
and/or decision sets, in this paper we explicitly focus on the scale-free
property. Regret of scale-free algorithms is proportional to the scale of the
losses, ensuring optimal linear dependency on the maximum norm of the loss
vectors.

We design and analyze three scale-free algorithms. First two algorithms, which
we call \textsc{AdaFTRL} and \textsc{SOLO FTRL}, are instances of
\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive learning
rate.  \textsc{AdaFTRL} is a generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  The third algorithm, which we call \textsc{Scale-Free Mirror
Descent}, is a variant of \textsc{Mirror Descent}. It is a generalization of
\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to arbitrary strongly convex
regularizers, similar to \textsc{Optimistic MD}~\cite{Rakhlin-Sridharan-2013}.
The three algorithms are presented and analyzed in
Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
\ref{section:mirror-descent} respectively.

>>>>

Online Linear Optimization (OLO) is a problem where an algorithm repeatedly
chooses a point $w_t$ from a convex decision set $K$, observes an arbitrary, or
even adversarially chosen, loss vector $\ell_t$ and suffers the loss $\langle
\ell_t, w_t \rangle$.  The goal of the algorithm is to have a small cumulative
loss. The performance of an algorithm is evaluated by the so-called regret, which
is the difference of the cumulative losses of the algorithm and of the
(hypothetical) strategy that would choose in every round the same best point in
hindsight.

@@

Rather than trying to design strategies for a particular form of loss vectors
and/or decision sets, in this paper we explicitly focus on the scale-free
property. The regret of scale-free algorithms is proportional to the scale of the
losses, ensuring optimal linear dependency on the maximum norm of the loss
vectors.

We design and analyze three scale-free algorithms. First two algorithms, which
we call \textsc{AdaFTRL} and \textsc{SOLO FTRL}, are instances of
\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive regularizers.
\textsc{AdaFTRL} is a generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  The third algorithm, which we call \textsc{Scale-Free Mirror
Descent}, is a variant of \textsc{Mirror Descent}. It is a generalization of
\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to arbitrary strongly convex
regularizers, similar to \textsc{Optimistic MD}~\cite{Rakhlin-Sridharan-2013}.
The three algorithms are presented and analyzed in
Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
\ref{section:mirror-descent} respectively.

### 2015-12-09 mirror-descent.tex
Reworked introduction
typos and small changes

\textsc{Mirror Descent}\footnote{The algorithm was originally proposed in
\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}.}
is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, Bregman
divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
\rangle$ needs to be defined. This is done by choosing a subgradient map $\grad
R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient of
$R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

@@

Generalizing existing regret bound for
\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} (see
also~\cite{Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}) we
can upper bound regret of \textsc{Scale-Free Mirror Descent}.  Proof is
in~\ref{section:mirror-descent-proofs}.

@@

We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
function and optimize $\lambda$. The result is the following Corollary. Its
proof is trivial.

@@

The regret bound for \textsc{Scale-Free Mirror Descent} in the
Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.  Similarly, regret bound in
Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and regret bounds in
Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

@@

Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
factors, regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
better and possibly worse than the regret bounds for \textsc{AdaFTRL} and
\textsc{SOLO FTRL}.  In fact, it is not hard to come up with examples where
$R(u)$ is finite whereas $\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention
two such examples. The first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined
on the whole space $V$, where for any $u \in V$, $R(u)$ is a finite value but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 =
+\infty$. The second example is the negative entropy regularizer $R(w) = \ln d
+ \sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex
$K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u
\in K$, $R(u)$ is finite and in fact lies in the interval $[-\ln d, 0]$ but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
\infty$. We return to these examples in the following subsection.

\subsection{Lower Bounds}

@@

The first counter-example is stated as
Theorem~\ref{theorem:first-counter-example} below. The decision set is the
whole space $K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note
that $R(w)$ is $1$-strongly convex with respect to $\|\cdot\|_2$ and the dual
norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. Corresponding Bregman divergence is
$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
the first vector of the standard unit basis.  On such sequence, Zinkevich's
\textsc{GIGA} and both versions of \textsc{AdaGrad with Composite Mirror
Update} have predictions identical to \textsc{Scale-Free Mirror Descent}. Hence
the lower bound applies to all four algorithms.

>>>>

\textsc{Mirror Descent}\footnote{The algorithm was originally proposed in
\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for convex optimization. In the online literature it makes the first appeareance, under a different name, in \cite{Kivinen-Warmuth-1997}.}
is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the Bregman
divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
\rangle$ needs to be defined. This is done by choosing a subgradient map $\grad
R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient of
$R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

@@

We can upper bound regret of \textsc{Scale-Free Mirror Descent} (see also \cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).  Proof is
in~\ref{section:mirror-descent-proofs}.

@@

We choose the regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
function and optimize $\lambda$. The result is the following Corollary. Its
proof is trivial.

@@

The regret bound for \textsc{Scale-Free Mirror Descent} in the
Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

@@

Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
factors, the regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
better and possibly worse than the regret bounds for \textsc{AdaFTRL} and
\textsc{SOLO FTRL}.  In fact, it is not hard to come up with examples where
$R(u)$ is finite whereas $\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention
two such examples. The first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined
on the whole space $V$, where for any $u \in V$, $R(u)$ is a finite value but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 =
+\infty$. The second example is the negative entropy regularizer $R(w) = \ln d
+ \sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex
$K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u
\in K$, $R(u)$ is finite and in fact lies in the interval $[-\ln d, 0]$ but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
\infty$. We return to these examples in the following subsection.

\subsection{Lower Bounds}
\label{subsection:mirror-descent-lower-bound}

@@

The first counter-example is stated as
Theorem~\ref{theorem:first-counter-example} below. The decision set is the
whole space $K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note
that $R(w)$ is $1$-strongly convex with respect to $\|\cdot\|_2$ and the dual
norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
the first vector of the standard unit basis.  On such sequence, Zinkevich's
\textsc{GIGA} and both versions of \textsc{AdaGrad with Composite Mirror
Update} have predictions identical to \textsc{Scale-Free Mirror Descent}. Hence
the lower bound applies to all four algorithms.

### 2015-12-09 per-coordinate.tex
typos and small changes

\section{Per-Coordinate Learning}

An interesting class of algorithms proposed in~\cite{McMahan-Streeter-2010} and
\cite{Duchi-Hazan-Singer-2011} are based on the so-called per-coordinate
learning rates.  As shown in \cite{Streeter-McMahan-2010}, our algorithms, or
in fact any algorithm for OLO, can be used with per-coordinate learning rates
as well.

Abstractly, we assume that the decision set is a Cartesian product $K=K_1
\times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
factor $K_j$, $j=1,2,\dots,d$, we can run any OLO algorithm separately and we
denote by $\Regret_T^{(j)}(u_j)$ its regret with respect to $u_j \in K_j$. The
overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
written as
$$
\Regret_T(u) = \sum_{j=1}^d \Regret_T^{(j)}(u_j) \; .
$$
If the algorithm for each factor is scale-free, the overall algorithm is
clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
each factor $K_j$, we generalize and improve existing regret bounds
\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with
per-coordinate learning rates.

>>>>

\subsection{Per-Coordinate Learning}

An interesting class of algorithms proposed in~\cite{McMahan-Streeter-2010} and
\cite{Duchi-Hazan-Singer-2011} are based on the so-called per-coordinate
learning rates.  As shown in \cite{Streeter-McMahan-2010}, any algorithm for OLO can be used with per-coordinate learning rates as well.

Abstractly, we assume that the decision set is a Cartesian product $K=K_1
\times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
factor $K_j$, $j=1,2,\dots,d$, we can run any OLO algorithm separately and we
denote by $\Regret_T^{(j)}(u_j)$ its regret with respect to $u_j \in K_j$. The
overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
written as
$$
\Regret_T(u) = \sum_{j=1}^d \Regret_T^{(j)}(u_j) \; .
$$

If the algorithm for each factor is scale-free, the overall algorithm is
clearly scale-free as well. Hence, even if not explicitly mentioned in the text, any algorithm we present can be trivially transformed in a per-coordinate version.

%Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
%each factor $K_j$, we generalize and improve existing regret bounds
%\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with
%per-coordinate learning rates.

### 2015-12-09 preliminaries.tex
typos and small changes

Our scale-free online learning algorithms are versions of the \textsc{Follow
The Regularized Leader}\footnote{The name \textsc{Follow The Regularized
Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}. Independently, the
algorithm was proposed by~\cite{Nestorov-2009} under the name \textsc{Dual
Averaging} and was later anlyzed in~\cite{Xiao-2010}.} (\textsc{FTRL})
algorithm with varying regularizers, presented as
Algorithm~\ref{algorithm:ftrl-varying-regularizer}. The algorithm is
paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
following lemma bounds its regret.

\begin{lemma}[Regret of Follow The Regularized Leader]
\label{lemma:generic-regret-bound}
For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
semi-continuous regularizers, regret of
Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}

Proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}.
For completeness, we include it in~\ref{section:definitions-proofs}.

>>>>

Our scale-free online learning algorithms are versions of the \textsc{Follow
The Regularized Leader}\footnote{The name \textsc{Follow The Regularized
Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}. It dates back to the potential-based forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was heavily developed by \cite{Shalev-Shwartz-2007}. 
Independently, the same algorithm was proposed by~\cite{Nestorov-2009} for convex optimization under the name \textsc{Dual
Averaging} and later rediscovered in~\cite{Xiao-2010} for online convex optimization. The version with time-varying regularizers was analyzed in \cite{Orabona-Crammer-Cesa-Bianchi-2014}.
} (\textsc{FTRL})
algorithm with \emph{varying regularizers}, presented as
Algorithm~\ref{algorithm:ftrl-varying-regularizer}. The algorithm is
paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
following lemma bounds its regret.

\begin{lemma}[Regret of Follow The Regularized Leader]
\label{lemma:generic-regret-bound}
For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
semi-continuous regularizers, the regret of
Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}

The proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}.
For completeness, we include it in~\ref{section:definitions-proofs}.

### 2015-12-09 solo-ftrl.tex
typos and small changes

The closest algorithm to a scale-free one in the OLO literature is the
\textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a
regularizer on each coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the
norm of the loss vectors. Identical considerations hold for the
\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
scale-free setting of the learning rate is valid only in the bounded case.

One simple approach would be to use a doubling trick on $\delta$ in order to
estimate on the fly the maximum norm of the losses. Note that a naive strategy
would still fail because the initial value of $\delta$ should be data-dependent
in order to have a scale-free algorithm. Moreover, we would have to upper bound
the regret in all the rounds where the norm of the current loss is bigger than
the estimate. Finally, the algorithm would depend on an additional parameter,
the ``doubling'' power. Hence, even guaranteeing a regret bound\footnote{For
lack of space, we cannot include the regret bound for the doubling trick
version. It would be exactly the same as in
Theorem~\ref{theorem:regret-solo-ftrl}, following a similar analysis, but with
the additional parameter of the doubling power.}, such strategy would give the
feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a
scale-free algorithm.

>>>>

The closest algorithm to a scale-free one in the OLO literature is the
\textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a
regularizer on each coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the
norm of the gradient vectors. Identical considerations hold for the
\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
scale-free setting of the learning rate is valid only in the bounded case.

One simple approach would be to use a doubling trick on $\delta$ in order to
estimate on the fly the maximum norm of the losses. Note that a naive strategy
would still fail because the initial value of $\delta$ should be data-dependent
in order to have a scale-free algorithm. Moreover, we would have to upper bound
the regret in all the rounds where the norm of the current loss is bigger than
the estimate. Finally, the algorithm would depend on an additional parameter,
the ``doubling'' power. Hence, even in the case one would prove a regret bound,
%\footnote{For
%lack of space, we cannot include the regret bound for the doubling trick
%version. It would be exactly the same as in
%Theorem~\ref{theorem:regret-solo-ftrl}, following a similar analysis, but with
%the additional parameter of the doubling power.},
such strategy would give the
feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a
scale-free algorithm.
